---
title: "Descriptive analysis"
author: "Johannes Zauner"
format: 
  html:
    self-contained: true
    code-tools: true
---

# Preface

This is a work-in-progress descriptive analysis of the `AkuffoEtAl2025` dataset.

```{r}
#| label: setup
#| include: false
library(LightLogR)
library(glue)
library(tidyverse)
library(gt)
library(cowplot)
library(legendry)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(patchwork)

source("https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/overview_plot.R")
```

# Overview

## Data import: wearable data

The first step is the import of wearable data from the `head` position (mounted on glasses).

```{r}
#| label: "general information"
#time zone of Kumasi
tz = "Africa/Accra"
#coordinates for Kumasi
coordinates <- c(6.6750074282377385, -1.572643823555129)
#regex to extract participant Id and wearing position
# pattern <- "[A-Z]+_S[0-9]{3}_[hcw]"
#regex to extract participant Id
pattern <- "[A-Z]+_S[0-9]{3}"

country_colors <- c(
  Sweden     = "#88CCEE",  # Sky blue
  Spain      = "#CC6677",  # Coral red
  Germany    = "#DDCC77",  # Mustard yellow
  Netherlands= "#117733",  # Dark green
  Turkey     = "#332288",  # Indigo
  Ghana      = "#AA4499",  # Purple-pink
  Costa_Rica = "#44AA99"   # Teal
)

```

```{r}
#path to participants
path_part1 <- "../data/raw/individual"
#path to actlumus data sans wearing position
path_part2 <- "/continuous/actlumus_"
#wearing position
wearing_position <- "head"
#getting all subfolders
folders <- dir(path_part1)
#creating complete folder names
paths <- glue("{path_part1}/{folders}{path_part2}{wearing_position}")
#collecting file names
files <- list.files(paths, full.names = TRUE)
files <- files[str_detect(files, "Report", negate = TRUE)]
```

```{r}
data <- import$ActLumus(files, tz, auto.id = pattern)
```

## Regularizing data

```{r}
#| fig-height: 15
#| fig-width: 5
data |> has_gaps()
data |> has_irregulars()
data |> gg_gaps(group.by.days = TRUE, show.irregulars = TRUE, full.days = FALSE)
```

Apparently, `KNUST_S003`, `KNUST_S004`, and `KNUST_S005` suffer from some kind of defect, where the data become irregular and show values around 10^4 lx, even at night. We first bin the irregular data into the next regular interval, and then convert implicit to explicit days.

```{r}
#| fig-height: 15
#| fig-width: 5
data_cleaned <- 
data |> 
  aggregate_Datetime() |> #remove irregular data
  gap_handler(full.days = TRUE) #remove implicit gaps

data_cleaned |> has_gaps() #no more implicit gaps
data_cleaned |> has_irregulars() #no more irregular data
data_cleaned |> gap_table()
```

# Dealing with measurement spikes

Measurement spikes seem to happen at random during certain periods of time. We can capture those by treating them as clusters of at maximum 2 epochs length. The next step visualizes this approach for `KNUST_S005`.

```{r}
#| fig-height: 12
#| fig-width: 8
data_cleaned |> 
  filter(Id == "KNUST_S005") |> 
  add_clusters(MEDI > 2000 & MEDI < 30000, 
                   cluster.duration = "20 secs", 
                   duration.type = "max") |>  
  gg_day(geom = "line") |> 
  gg_photoperiod(coordinates) +
  geom_point(data = \(x) filter(x, !is.na(state)), col = "red") +
  labs(title = "Identification of sudden measurement spikes (red dots)")
```

We assess how many there are across the participants:

```{r}
data_cleaned |> 
  extract_clusters(MEDI > 2000 & MEDI < 30000, 
                   cluster.duration = "20 secs", 
                   duration.type = "max",
                   add.label = TRUE) |> 
  group_by(label, Id) |> 
  summarize_numeric() |> 
  select(Id, label, episodes) |> 
  group_by(label) |> 
  gt() |> 
  tab_header("Measurement spikes")
```

Spikes are detected for every Id. We will not remove the data, but rather replace them with the value 2 intervals prior. This ensures plausible values for that time and avoids that a spike is copied over.

```{r}
#| fig-height: 6
#| fig-width: 12
#| warning: false
data_cleaned <-
  data_cleaned |>
  add_clusters(MEDI > 2000 & MEDI < 30000,
               cluster.duration = "20 secs",
               duration.type = "max") |>
  mutate(MEDI =
           ifelse(!is.na(state), lag(MEDI, 2), MEDI))

data_cleaned |> 
  filter(Id %in% c("KNUST_S003", "KNUST_S004", "KNUST_S005")) |> 
  gg_days()
```

These data now look sensible enough for further visualization and processing.

## Visualization

### Overview

```{r}
#| warning: false
#| message: false
#| fig-height: 8
#| fig-width: 10

data_cleaned |> 
  mutate(Id = Id |> fct_relabel(\(x) str_remove(x, "KNUST_"))) |> 
grand_overview(coordinates, "Kumasi", "Ghana", country_colors, photoperiod_sequence = 0.1)

ggsave("../output/figures/Figure_1.png", width = 17, height = 10, scale = 2, units = "cm")
```



