---
title: "Descriptive analysis"
author: "Johannes Zauner"
format: 
  html:
    self-contained: true
    code-tools: true
---

# Preface

This is a work-in-progress descriptive analysis of the `AkuffoEtAl2025` dataset.

```{r}
#| label: setup
#| include: false
library(LightLogR)
library(glue)
library(tidyverse)
library(here)
library(gt)
library(cowplot)
library(legendry)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(patchwork)

source(
  "https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/overview_plot.R"
)
source(
  "https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/summary_table.R"
)
```

# Overview

## Data import: wearable data

The first step is the import of wearable data from the `head` position (mounted on glasses).

```{r}
#| label: "general information"
#time zone of Kumasi
tz = "Africa/Accra"
#coordinates for Kumasi
coordinates <- c(6.6750074282377385, -1.572643823555129)
#regex to extract participant Id and wearing position
# pattern <- "[A-Z]+_S[0-9]{3}_[hcw]"
#regex to extract participant Id
pattern <- "[A-Z]+_S[0-9]{3}"

country_colors <- c(
  Sweden = "#88CCEE", # Sky blue
  Spain = "#CC6677", # Coral red
  Germany = "#DDCC77", # Mustard yellow
  Netherlands = "#117733", # Dark green
  Turkey = "#332288", # Indigo
  Ghana = "#AA4499", # Purple-pink
  Costa_Rica = "#44AA99" # Teal
)

```

```{r}
#path to participants
path_part1 <- here("data/raw/individual")
#path to actlumus data sans wearing position
path_part2 <- "/continuous/actlumus_"
#wearing position
wearing_position <- "head"
#getting all subfolders
folders <- dir(path_part1)
#creating complete folder names
paths <- glue("{path_part1}/{folders}{path_part2}{wearing_position}")
#collecting file names
files <- list.files(paths, full.names = TRUE)
files <- files[str_detect(files, "Report", negate = TRUE)]
```

```{r}
data <- import$ActLumus(files, tz, auto.id = pattern)
```

## Regularizing data

```{r}
#| fig-height: 15
#| fig-width: 5
data |> has_gaps()
data |> has_irregulars()
data |> gg_gaps(group.by.days = TRUE, show.irregulars = TRUE, full.days = FALSE)
```

Apparently, `KNUST_S003`, `KNUST_S004`, and `KNUST_S005` suffer from some kind of defect, where the data become irregular and show values around 10^4 lx, even at night. We first bin the irregular data into the next regular interval, and then convert implicit to explicit days. We will further remove unreasonably high values above 100.000 lx melanopic EDI

```{r}
#| fig-height: 15
#| fig-width: 5
data_cleaned <-
  data |>
  aggregate_Datetime() |> #remove irregular data
  filter(MEDI <= 10^5) |> #remove high values
  gap_handler(full.days = TRUE) #remove implicit gaps

data_cleaned |> has_gaps() #no more implicit gaps
data_cleaned |> has_irregulars() #no more irregular data
data_cleaned |> gap_table()
```

# Dealing with measurement spikes

Measurement spikes seem to happen at random during certain periods of time. We can capture those by treating them as clusters of at maximum 2 epochs length. The next step visualizes this approach for `KNUST_S005`. 

```{r}
#| fig-height: 12
#| fig-width: 8
data_cleaned |>
  filter(Id == "KNUST_S005") |>
  add_clusters(
    MEDI > 2000 & MEDI < 30000,
    cluster.duration = "20 secs",
    duration.type = "max"
  ) |>
  gg_day(geom = "line") |>
  gg_photoperiod(coordinates) +
  geom_point(data = \(x) filter(x, !is.na(state)), col = "red") +
  labs(title = "Identification of sudden measurement spikes (red dots)")
```

We assess how many there are across the participants:

```{r}
data_cleaned |>
  extract_clusters(
    MEDI > 2000 & MEDI < 30000,
    cluster.duration = "20 secs",
    duration.type = "max",
    add.label = TRUE
  ) |>
  group_by(label, Id) |>
  summarize_numeric() |>
  select(Id, label, episodes) |>
  group_by(label) |>
  gt() |>
  tab_header("Measurement spikes")
```

Spikes are detected for every Id. We will not remove the data, but rather replace them with the value 2 intervals prior. This ensures plausible values for that time and avoids that a spike is copied over.

```{r}
#| fig-height: 6
#| fig-width: 12
#| warning: false
data_cleaned <-
  data_cleaned |>
  add_clusters(
    MEDI > 2000 & MEDI < 30000,
    cluster.duration = "20 secs",
    duration.type = "max"
  ) |>
  mutate(MEDI = ifelse(!is.na(state), lag(MEDI, 2), MEDI))

data_cleaned |>
  filter(Id %in% c("KNUST_S003", "KNUST_S004", "KNUST_S005")) |>
  gg_days()
```

These data now look sensible enough for further visualization and processing.

## Single out extreme patterns

```{r}
#| fig-width: 18
#| fig-height: 35
data_2min <- 
data_cleaned |> 
  select(Id, Datetime, MEDI) |> 
  aggregate_Datetime("5 mins", numeric.handler = \(x) mean(x, na.rm = TRUE)) |> 
  remove_partial_data(
    MEDI, by.date = TRUE,
  ) |> 
  mutate(.Date = (date(Datetime) - min(date(Datetime)) + 1))
  
dose_data <- 
data_2min |> 
  add_Date_col(group.by = TRUE) |> 
  summarize(
    dose(
      MEDI,
      Datetime,
      na.rm = TRUE,
      as.df = TRUE
    ),
    .groups = "drop"
  )

dose_data |> 
  ungroup() |> 
  filter(dose == max(dose) | dose == min(dose))

data_2min |> 
  gg_day(facetting = FALSE, 
         group = factor(.Date),
         geom = "line", 
         aes_col = factor(.Date)) +
  facet_wrap(~Id, ncol = 2)

data_2min |>
  # filter(Id %in% c("UCR_S007", "UCR_S001")) |>
  filter(Id %in% c("KNUST_S007")) |>
  # filter(date(Datetime) %in% c("2025-07-01", "2025-07-05")) |>
  add_Date_col(group.by = TRUE) |>
  gg_doubleplot(aes_col = factor(.Date), aes_fill = factor(.Date)) +
  guides(fill = "none", color = "none")

```

## Exporting hourly values

```{r}
data_export <- 
data_cleaned |> 
  aggregate_Datetime("1 minute", numeric.handler = \(x) mean(x, na.rm = TRUE)) |> 
  remove_partial_data(MEDI, threshold.missing = "3 hours", by.date = TRUE)

save(data_export, file = "../data/gh_1minute.RData")
```

## Visualization

### Overview

```{r}
#| warning: false
#| message: false

data_cleaned |>
  mutate(Id = Id |> fct_relabel(\(x) str_remove(x, "KNUST_"))) |>
  grand_overview(
    coordinates,
    "Kumasi",
    "Ghana",
    country_colors,
    photoperiod_sequence = 0.1
  )

ggsave(
  here("output/figures/Figure_1.png"),
  width = 17,
  height = 10,
  scale = 2,
  units = "cm"
)
ggsave(
  here("output/figures/Figure_1.jpeg"),
  width = 17,
  height = 10,
  scale = 2,
  units = "cm"
)
```

## Stats

### Summary table

```{r}
#| message: false
#| warning: false

table_summary <-
light_summary_table(
  data_cleaned, coordinates, "Kumasi", "Ghana", country_colors["Ghana"],
  histograms = TRUE
)

table_summary

gtsave(table_summary, here("output/tables/table_summary.png"), vwidth = 850)
gtsave(table_summary, here("output/tables/table_summary.pdf"))
# gtsave(table_summary |> cols_hide(c(plot, footnote)), here("output/tables/table_summary.docx"))


```
