---
title: "Import mHLEA"
author: "Johannes Zauner"
format: 
  html:
    self-contained: true
    code-tools: true
---

## Preface

This document imports the `mHLEA` evening diary (light exposure) and shows descriptive statistics for the site.

## Setup

```{r}
#| message: false
library(tidyverse)
library(LightLogR)
library(glue)
library(readxl)
library(gt)
library(gtsummary)
source("../project_globals.R")
remote <- 
  "https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/"

c("labeling",
  "radio_factors",
  "time_summaries",
  "prepare_codebook",
  "filefinder",
  "add_label",
  "who5_scoring",
  "general_parameters",
  "coltype_checker",
  "diarydate",
  "tables"
) |> walk(\(x) source(paste0(remote, x, ".R")))
```

## Preparation

```{r}
#collect codebook
codebook <- prepare_codebook("MeLiDosEveningDiaries_DataDictionary_2024-10-16.csv", 
                             form.filter = c("light_exposure_diary", "form_1"))
#collect files
files <- filefinder("mHLEA_digital", continuous = TRUE, individual = TRUE)
#import files
data <-   read_csv(files[c(1:2,4:15)], show_col_types = FALSE) |> 
  bind_rows(read_csv(files[c(3)], show_col_types = FALSE)) |> 
  drop_na(redcap_repeat_instance) |> 
  arrange(record_id, startdate) |> 
  mutate(record_id = sprintf("KNUST_S%03d", record_id))
#check column types
coltype_check <- coltype_checker(codebook, data)
coltype_check$details |> gt()
#collect relevant columns: POSIXct, Date & numeric
relevant_columns <- 
  coltype_check$details |> 
  pull(col) |> 
  union("startdate")
#select relevant columns
data <- data |> select(any_of(relevant_columns))
#label variables
data <-
data |> 
  add_radio_factors(codebook, 
                    var_col = `Variable / Field Name`, 
                    type_col = `Field Type`,
                    levels_col = `Choices, Calculations, OR Slider Labels`
                    ) |> 
  add_col_labels(codebook, var_col = `Variable / Field Name`, label_col = `Field Label`) |> 
  select(record_id, startdate, mhlea_confidence)
```

## Set relevant dates

```{r}
#if data was collected between 14:00 and 24:00, it is assigned to the same day.
#if collected between 00:00 and 13:59, it is assigned to the previous day.

data <- data |> diarydate(startdate)
attr(data$Date, "label") <- "Date"
```

## Import paper diaries

```{r}
#collect files
files <- 
  filefinder("mHLEA_paper", continuous = TRUE, individual = TRUE)
file_ids <- files |> basename() |> str_extract(paste0(site, "_S.{3}"))
files <- files |> set_names(file_ids)
#import files
data_paper <-
  map(files, read_excel) |> 
    map(\(x) if(nrow(x) > 0) x |> 
          mutate(activity = as.character(activity))
        )
data_paper <-
  data_paper[which(!map_lgl(data_paper, is.null))] |> 
  list_rbind(names_to = "record_id") |> 
  mutate(timestamp = 
           round_date(timestamp, unit = "hour") |> 
           force_tz(tzs[[site]]),
         Date = date(timestamp),
         start = timestamp,
         end = timestamp + dhours(1),
         .before = timestamp) |> 
  select(-timestamp)
```

## Combine paper diaries with confidence rating

```{r}
data <- 
data_paper |> 
  left_join(data, by = c("record_id", "Date")) |> 
  relocate(startdate, .after = last_col())
```

## Translate comments into english

Native language is translated into English with AI and later checked by a site researcher.

```{r}
#| label: translate into english with AI
# 
# library(ellmer)
# 
# #Providing the relevant codebook portions
# chat <- chat_openai(paste0("Clean the dataset according to the instructions in the output structure."))
# 
# #Providing the input
# data_red1 <-
# data|>
#   select(record_id, start, activity_desc) |>
#   filter(!is.na(activity_desc))
# 
# data_red <- data_red1 |>
#   pmap(~ paste(paste(names(data_red1), c(...), sep = ": "), collapse = ", "))
# 
# #creating an output structure
# type_data <- type_object(
#   record_id = type_string("copy the record_id information"),
#   start = type_string("copy the start information as 'YYYY-MM-DD HH:SS:MM'"),
#   activity_desc_english = type_string("Translate (or if in english already, copy) the 'activity_desc' column here. Question: Please specify your activity.", required = FALSE),
# )
# 
# data_llm <-
# parallel_chat_structured(
#   chat,
#   data_red,
#   type = type_data
# )
# 
# #Ensure that no NA is caught as string
# data_llm <-
#   data_llm |>
#   mutate(across(everything(), \(x) case_when(x == "NA" ~ NA, .default = x)))
# 
# #check that input and output are identical
# stopifnot("Input must by identical to output check" =
#             all(data_llm$record_id == data_red$record_id, na.rm = TRUE))
# stopifnot("Input must by identical to output check" =
#             all(data_llm$start == data_red$start, na.rm = TRUE))
# 
# data_llm$start <- data_red1$start
# 
# data_llm$activity_desc <- data_red1$activity_desc
# 
# path <- "../data/AI_translations/"
# if(!dir.exists(path)) dir.create(path, recursive = TRUE)
# write_csv(data_llm, "../data/AI_translations/lightexposurediary.csv")

```

```{r}
data_llm <- 
  read_csv("../data/AI_translations/lightexposurediary.csv") |> 
  select(activity_desc, activity_desc_english) |> 
  distinct(activity_desc, .keep_all = TRUE)
#add output to original
data <- data |> left_join(data_llm, by = c("activity_desc"))
data <-
  data |> 
  mutate(across(ends_with("english"), \(x) add_label(x, "Please specify your activity (English translation)")
                ))
```

## Dissecting answers

```{r}
data <- 
data |> 
  filter(!is.na(lightsource) | !is.na(activity)) |> 
  mutate(lightsource = str_remove_all(lightsource, "[^[:alpha:]]"),
         light_electric_indoor = str_detect(lightsource, "L|l"),
         light_electric_outdoor = str_detect(lightsource, "S|s"),
         light_daylight_indoor = str_detect(lightsource, "I|i"),
         light_daylight_outdoor = str_detect(lightsource, "O|o"),
         light_display = str_detect(lightsource, "E|e"),
         light_sleep_darkness = str_detect(lightsource, "D|d"),
         light_sleep_imission = str_detect(lightsource, "W|w"),
         .after = lightsource) |> 
  separate_wider_position(lightsource, 
                          widths = c("primary" = 1, "secondary" = 1),
                          names_sep = "_",
                          too_few = "align_start",
                          too_many = "drop",
           ) |> 
  mutate(activity = str_remove_all(activity, "[^[:digit:]]"),
         act_sleep = str_detect(activity, "1"),
         act_home = str_detect(activity, "2"),
         act_road_vehicle = str_detect(activity, "3"),
         act_road_open = str_detect(activity, "4"),
         act_working_indoor = str_detect(activity, "5"),
         act_working_outdoor = str_detect(activity, "6"),
         act_free_outdoor = str_detect(activity, "7"),
         act_other = str_detect(activity, "8"),
         .after = activity
         ) |> 
  select(-activity)
  
```

Adding labels:

```{r}
factor_levels_mHLEA <- c(
    L = "Electric light source indoors",
    S = "Electric light source outdoors",
    I = "Daylight indoors",
    O = "Daylight outdoors (including shade)",
    E = "Emissive display light",
    D = "Darkness during sleep",
    W = "Light entering from outside during sleep"
)
data <-
data |> 
  mutate(
    across(
      contains("lightsource"), 
               \(x) x |> 
                    str_to_upper() |> 
                    factor( 
                           levels = names(factor_levels_mHLEA),
                           labels = factor_levels_mHLEA
               )
      )
    )

```

```{r}
labels_mHLEA <- 
  c(
    record_id = "Record ID",
    Date = "Date",
    start = "Beginning timestamp",
    end = "Ending timestamp",
    lightsource_primary = "Primary lightsource (mH-LEA)",
    lightsource_secondary = "Secondary lightsource (mH-LEA)",
    light_electric_indoor = "Electric light source indoors (mH-LEA)",
    light_electric_outdoor = "Electric light source outdoors (mH-LEA)",
    light_daylight_indoor = "Daylight indoors (mH-LEA)",
    light_daylight_outdoor = "Daylight outdoors (including shade) (mH-LEA)",
    light_display = "Emissive display light (mH-LEA)",
    light_sleep_darkness = "Darkness during sleep (mH-LEA)",
    light_sleep_imission = "Light entering from outside during sleep (mH-LEA)",
    act_sleep = "Sleeping in bed (activity)",
    act_home = "Awake at home (activity)",
    act_road_vehicle = "On the road with public transport/car (activity)",
    act_road_open = "On the road with bike/on foot (activity)",
    act_working_indoor = "Working in the office/from home (activity)",
    act_working_outdoor = "Working outdoors (including lunch break outdoors) (activity)",
    act_free_outdoor = "Free time outdoors (activity)",
    act_other = "Other (activity)",
    act_desc = "Other activity description",
    act_desc_english = "Other activity (english translation)",
    startdate = "Starting time to fill in questionnaire"
  )

data <- add_labels(data, labels_mHLEA)
```

## Summarize results

```{r}
table <- 
  data |> 
  select(-c(start, end, Date, activity_desc, activity_desc_english, startdate)) |> 
  table_general("Light exposure (mH-LEA) and activity diary")
table
gtsave(table |> as_gt(), filename = "../output/tables/table_lightexposurediary.png", vwidth = 800)
```

### Export

```{r}
data <- data |> rename(Id = record_id)
lightexposurediary <- data
path <- "../data/imported/continuous/"
if(!dir.exists(path)) dir.create(path, recursive = TRUE)
save(lightexposurediary, file = "../data/imported/continuous/lightexposurediary.RData")
```
