---
title: "Import chronotype"
author: "Johannes Zauner"
format: 
  html:
    self-contained: true
    code-tools: true
editor_options: 
  chunk_output_type: console
---

## Preface

This document imports the `chronotype` and shows descriptive statistics for the site.

## Setup

```{r}
#| message: false
library(tidyverse)
library(LightLogR)
library(glue)
library(readxl)
library(gt)
library(gtsummary)
library(mctq)

remote <- 
  "https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/"

c("labeling",
  "radio_factors",
  "prepare_codebook",
  "time_summaries",
  "filefinder",
  "general_parameters",
  "coltype_checker",
  "tables"
) |> walk(\(x) source(paste0(remote, x, ".R")))
```

#calculate mctq

```{r}
#collect codebook
codebook <- 
  prepare_codebook("MeLiDosIntakeQuestionnaires_DataDictionary_2024-10-16.csv")
#collect files
files <- filefinder("chronotype",individual = FALSE, negate = "lookup|meq")
#import files
data <- read_csv(files, show_col_types = FALSE)
#check column types
coltype_check <- coltype_checker(codebook, data)
coltype_check$details |> gt()
#collect relevant columns
relevant_columns <- 
  coltype_check$details |> 
  pull(col)
#select relevant columns
data <- data |> select(any_of(relevant_columns))
data <- 
  data |> 
  mutate(record_id = sprintf("KNUST_S%03d", record_id))
```

2 participants provided an unlikely bedtime before noon (and have not indicated shiftwork) - these participants will be treated as if their bedtime is 12 hours later (problem with am/pm)

```{r}
Id_to_shift <- c("KNUST_S001", "KNUST_S013")

data <-
  data %>% 
  mutate(across(contains(c("ready_sleep_", "bedtime")), 
                \(x) case_when(
                  record_id %in% Id_to_shift ~ (x + hms::hms(hours = 12)) %>% hms::as_hms(),
                  .default = x))
         )
```

3 participants provided an unlikely early end of work - these will be increased by 12 hours

```{r}
Id_to_shift <- c("KNUST_S001", "KNUST_S009", "KNUST_S011")

data <-
  data %>% 
  mutate(across(contains(c("work_end")), 
                \(x) case_when(
                  record_id %in% Id_to_shift ~ (x + hms::hms(hours = 12)) %>% hms::as_hms(),
                  .default = x))
         )
```

Furthermore, some participants indicated that they are ready to sleep before going to bed. In the way the questionnaire is analysed, this has to be after they go to bed. Thus we will take the later time in both cases (i.e., if bedtime is later then ready to sleep, we will use this).

```{r}
data <-
 data %>%
 pivot_longer(
    cols = contains(c("ready_sleep_", "bedtime")),
    names_to = c(".value", "context"),
    names_pattern = "^(mctq_bedtime|mctq_ready_sleep)_(.*)$"
  ) %>%
  mutate(
    mctq_ready_sleep = pmax(mctq_bedtime, mctq_ready_sleep),
    mctq_nr_workdays_adj = ifelse(is.na(mctq_nr_workdays), 5, mctq_nr_workdays)
  ) %>%
  pivot_wider(names_from = context, values_from = c(mctq_bedtime, mctq_ready_sleep))
```

> Note: participants that indicated no number of regular work days, but that still provided numbers for work- and free days will be calculated as 5 work days and 2 free days.

```{r}
#reorder variables, remove mctq variable, recode falling asleep to minutes
data <- 
  data |> 
  relocate(any_of(codebook$`Variable / Field Name`)) |> 
  rename_with(\(x) str_remove(x, "mctq_")) |> 
  mutate(across(contains("fall_sleep"), ~as.duration(.x*60)))

#label variables
data <-
data |> 
  add_radio_factors(codebook |> mutate(`Variable / Field Name` = `Variable / Field Name` |> str_remove("mctq_")), 
                    var_col = `Variable / Field Name`, 
                    type_col = `Field Type`,
                    levels_col = `Choices, Calculations, OR Slider Labels`
                    ) |> 
  add_col_labels(codebook |> mutate(`Variable / Field Name` = `Variable / Field Name` |> str_remove("mctq_")), 
                 var_col = `Variable / Field Name`, label_col = `Field Label`)
```

```{r}
data <- 
data %>% 
mutate(
    so_w = so(ready_sleep_work, fall_sleep_work),
    so_f = so(ready_sleep_free, fall_sleep_free),
    sd_w = sdu(so_w, wake_time_work),
    sd_f = sdu(so_f, wake_time_free),
    msw = msl(so_w, sd_w),
    msf = msl(so_f, sd_f),
    sd_week = sd_week(sd_w, sd_f, nr_workdays_adj),
    msf_sc = msf_sc(msf, sd_w, sd_f, sd_week, alarm_f = rep(FALSE, n())),
    sjl = sjl(msw, msf, abs = FALSE),
  .before = 2
    )

chronotype <- data |> select(record_id:sjl)
```
> Note: Participants who indicated they use an alarm clock on free days (alarm_free) cannot be used for MCTQ calculation. In this overview, this behavior is overwritten by setting the alarm_f parameter to FALSE. For formal analysis, these must be removed.

## meq

```{r}
#collect files
files <- filefinder("chronotype",individual = FALSE, negate = "lookup|mctq")
#import files
data <- read_csv(files, show_col_types = FALSE) |>
    drop_na(redcap_repeat_instance)
#check column types
coltype_check <- coltype_checker(codebook, data)
coltype_check$details |> gt()
#collect relevant columns
#select relevant columns
data <- data |> select(-redcap_repeat_instance, -redcap_repeat_instrument)
data <- 
  data |> 
  mutate(record_id = sprintf("KNUST_S%03d", record_id))
#label variables
data <-
data |> 
  add_radio_factors(codebook |> mutate(`Variable / Field Name` = `Variable / Field Name` |> str_remove("mctq_")), 
                    var_col = `Variable / Field Name`, 
                    type_col = `Field Type`,
                    levels_col = `Choices, Calculations, OR Slider Labels`
                    ) |> 
  add_col_labels(codebook |> mutate(`Variable / Field Name` = `Variable / Field Name` |> str_remove("mctq_")), 
                 var_col = `Variable / Field Name`, label_col = `Field Label`)
```

```{r}
data <- 
  data |> 
  #invert scores
  mutate(across(c("meq_freewake", "meq_freesleep", "meq_peak_perf", "meq_perceived_chron"),
         \(x) {
           n_levels <- levels(x) |> length()
           n_levels - as.numeric(x)
         }, .names = "{.col}_scored"),
         across(c("meq_alarm", "meq_free_bedtime", "meq_mor_exercise", "meq_eve_tired", "meq_delayed_sleep", "meq_phys_work", "meq_ideal_week", "meq_besthour"),
         \(x) {
           n_levels <- levels(x) |> length()
           n_levels - as.numeric(x) + 1
         }, .names = "{.col}_scored"),
         across(c("meq_wakeupease", "meq_mor_alert", "meq_mor_hunger", "meq_mor_feel", "meq_shift_conseq", "meq_eve_exercise"),
         \(x) {
           as.numeric(x)
         }, .names = "{.col}_scored"),
         meq_peak_perf_scored = meq_peak_perf_scored*2,
         meq_eleven_tired_scored = case_match(meq_eleven_tired,
                                             "Not at all tired" ~ 0,
                                             "A little tired" ~ 2,
                                             "Fairly tired" ~ 3,
                                             "Very tired" ~ 5),
         meq_perceived_chron_scored = meq_perceived_chron_scored*2
         )


meq_types <- c("Definitely morning type", "Moderately morning type",
               "Intermediate", "Moderately evening type",
               "Definitely evening type")

data <- 
  data |> 
  rowwise() |> 
  mutate(meq = sum(c_across(contains("_scored"))),
         meq_type = case_when(meq >= 70 ~ "Definitely morning type",
                              meq >= 59 ~ "Moderately morning type",
                              meq >= 42 ~ "Intermediate",
                              meq >= 31 ~ "Moderately evening type",
                              meq >= 16 ~ "Definitely evening type",
                              .default = NA) |> 
           factor(levels = meq_types)) |> 
  select(-contains("_scored"), -morning_eveningness_questionnaire_meq_complete)

attr(data$meq, "label") <- "Morningness-Eveningness-Score (calculated)"
attr(data$meq_type, "label") <- "MEQ chronotypes (calculated)"

```

## Summarize results

```{r}
chronotype <- left_join(chronotype, data |> select(record_id, meq, meq_type), by = "record_id")
chronotype <- 
  chronotype |> mutate(sjl = sjl |> as.duration(),
                       across(c(so_w, so_f, msw, msf, msf_sc), \(x) paste("2000-01-01", x) |> parse_datetime()))
attr(chronotype$so_w, "label") <- "Sleep onset on working days (calculated)"
attr(chronotype$so_f, "label") <- "Sleep onset on free days (calculated)"
attr(chronotype$sd_w, "label") <- "Sleep duration on working days (calculated)"
attr(chronotype$sd_f, "label") <- "Sleep duration on free days (calculated)"
attr(chronotype$msw, "label") <- "Mid-sleep on working days (calculated)"
attr(chronotype$msf, "label") <- "Mid-sleep on free days (calculated)"
attr(chronotype$sd_week, "label") <- "Average weekly sleep duration (calculated)"
attr(chronotype$msf_sc, "label") <- "Mid-sleep on free days, sleep corrected (calculated)"
attr(chronotype$sjl, "label") <- "Social jetlag (calculated)"

chronotype <- chronotype |> relocate(meq_type, meq, msf_sc, sjl, .after = record_id)

table_chronotype(chronotype)

gtsave(table_chronotype(chronotype) |> as_gt(), filename = "../output/tables/table_chronotype.png", vwidth = 800)
```

### Export

```{r}
chronotype <- 
  chronotype |> mutate(across(c(so_w, so_f, msw, msf, msf_sc), \(x) x |> hms::as_hms()))

attr(chronotype$so_w, "label") <- "Sleep onset on working days (calculated)"
attr(chronotype$so_f, "label") <- "Sleep onset on free days (calculated)"
attr(chronotype$msw, "label") <- "Mid-sleep on working days (calculated)"
attr(chronotype$msf, "label") <- "Mid-sleep on free days (calculated)"
attr(chronotype$msf_sc, "label") <- "Mid-sleep on free days, sleep corrected (calculated)"

chronotype <- chronotype |> rename(Id = record_id)
path <- "../data/imported/"
if(!dir.exists(path)) dir.create(path, recursive = TRUE)
save(chronotype, file = "../data/imported/chronotype.RData")


```
