---
title: "Import light head/glasses"
author: "Johannes Zauner"
format: 
  html:
    self-contained: true
    code-tools: true
---

# Preface

This is a work-in-progress descriptive analysis of the `AkuffoEtAl2025` dataset.

```{r}
#| label: setup
#| include: false
library(LightLogR)
library(glue)
library(tidyverse)
library(gt)
library(readxl)
library(cowplot)
library(legendry)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(patchwork)
library(rlang)
library(here)

remote <- 
  "https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/"

c(
  "filefinder",
  "general_parameters",
  "overview_plot"
) |> walk(\(x) source(paste0(remote, x, ".R")))

site <- "KNUST"
```

# Overview

## Data import: wearable data

The first step is the import of wearable data from the `head` position (mounted on glasses).

```{r}
#| label: "general information"
#regex to extract participant Id and wearing position
# pattern <- "[A-Z]+_S[0-9]{3}_[hcw]"
#regex to extract participant Id
pattern <- "[A-Z]+_S[0-9]{3}"
```

```{r}
files <- filefinder("actlumus_head", continuous = TRUE, negate = "Report")
```

```{r}
data <- import$ActLumus(files, tzs[site], auto.id = pattern)
```

## Regularizing data

In the first step, we will trim the data by the study time.

```{r}
path_study_dates <- paste0("../data/Study_dates_MeLiDos_", site, ".xlsx")

#import table with study times
Study_dates <- read_excel(path_study_dates)
#gather the important information
Study_dates <-
  Study_dates |> 
    rename(Id = subjectID_device, start = datetime_trial_start, end = datetime_trial_end) |> 
    select(Id, start, end) |> 
    mutate(across(c(start, end), \(x) force_tz(x, tzs[site])),
           trial = TRUE) |> 
    filter(str_detect(Id, "_h$")) |>
    mutate(Id = str_remove(Id, "_h$")) |> 
  group_by(Id)

#add the trim information to the dataset and filter by it
data <- 
  data |> 
  add_states(Study_dates) |> 
  dplyr::filter(trial) |> 
  select(-trial)

data |> gg_overview()
```

## Regularizing data

```{r}
#| fig-height: 15
#| fig-width: 5
data |> has_gaps()
data |> has_irregulars()
data |> gg_gaps(group.by.days = TRUE, show.irregulars = TRUE, full.days = FALSE)
```

Apparently, `KNUST_S003`, `KNUST_S004`, and `KNUST_S005` suffer from some kind of defect, where the data become irregular and show values around 10^4 lx, even at night. We first bin the irregular data into the next regular interval, and then convert implicit to explicit days. We will further remove unreasonably high values above 100.000 lx melanopic EDI

```{r}
#| fig-height: 15
#| fig-width: 5
data_cleaned <-
  data |>
  aggregate_Datetime() |> #remove irregular data
  filter(MEDI <= 10^5) |> #remove high values
  gap_handler(full.days = TRUE) #remove implicit gaps

data_cleaned |> has_gaps() #no more implicit gaps
data_cleaned |> has_irregulars() #no more irregular data
data_cleaned |> gap_table()
```

# Dealing with measurement spikes

Measurement spikes seem to happen at random during certain periods of time. We can capture those by treating them as clusters of at maximum 2 epochs length. The next step visualizes this approach for `KNUST_S005`. 

```{r}
#| fig-height: 12
#| fig-width: 8
data_cleaned |>
  filter(Id == "KNUST_S005") |>
  add_clusters(
    MEDI > 2000 & MEDI < 30000,
    cluster.duration = "20 secs",
    duration.type = "max"
  ) |>
  gg_day(geom = "line") |>
  gg_photoperiod(coordinates[[site]]) +
  geom_point(data = \(x) filter(x, !is.na(state)), col = "red") +
  labs(title = "Identification of sudden measurement spikes (red dots)")
```

We assess how many there are across the participants:

```{r}
data_cleaned |>
  extract_clusters(
    MEDI > 2000 & MEDI < 30000,
    cluster.duration = "20 secs",
    duration.type = "max",
    add.label = TRUE
  ) |>
  group_by(label, Id) |>
  summarize_numeric() |>
  select(Id, label, episodes) |>
  group_by(label) |>
  gt() |>
  tab_header("Measurement spikes")
```

Spikes are detected for every Id. We will not remove the data, but rather replace them with the value 2 intervals prior. This ensures plausible values for that time and avoids that a spike is copied over.

```{r}
#| fig-height: 6
#| fig-width: 12
#| warning: false
data_cleaned <-
  data_cleaned |>
  add_clusters(
    MEDI > 2000 & MEDI < 30000,
    cluster.duration = "20 secs",
    duration.type = "max"
  ) |>
  mutate(MEDI = ifelse(!is.na(state), lag(MEDI, 2), MEDI))

data_cleaned |>
  filter(Id %in% c("KNUST_S003", "KNUST_S004", "KNUST_S005")) |>
  gg_days()
```

These data now look sensible enough for further visualization and processing.

```{r}
#| fig-height: 15
#| fig-width: 5
data_cleaned <- 
data |> 
  cut_Datetime("10 secs", New.colname = Datetime) |> #remove irregular data
  gap_handler(full.days = TRUE) #remove implicit gaps

data_cleaned |> has_gaps() #no more implicit gaps
data_cleaned |> has_irregulars() #no more irregular data
data_cleaned |> gap_table(MEDI) |> cols_hide(ends_with("_n"))
```

## Exporting values

```{r}
data_cleaned <- 
  data_cleaned |> mutate(position = "glasses")

light_glasses_1min <- 
data_cleaned |> 
  aggregate_Datetime("1 minute", numeric.handler = \(x) mean(x, na.rm = TRUE)) |> 
  remove_partial_data(MEDI, threshold.missing = "3 hours", by.date = TRUE)

light_glasses <- data_cleaned

save(light_glasses_1min, file = "../data/imported/light/light_glasses_1minute.RData")
save(light_glasses, file = "../data/imported/light/light_glasses.RData")
```

## Visualization

```{r}
#| warning: false
#| message: false
#| fig-height: 8
#| fig-width: 12

prefix <- paste0(site, "_")

data_cleaned |> 
  mutate(Id = Id |> fct_relabel(\(x) str_remove(x, prefix))) |> 
grand_overview(coordinates[[site]], cities[[site]], countries[[site]], 
               country_colors[[site]], photoperiod_sequence = 0.1)

ggsave("../output/figures/Figure_1.png", width = 17, height = 10, scale = 2, units = "cm")
ggsave("../output/figures/Figure_1.jpeg", width = 17, height = 10, scale = 2, units = "cm")
```

## Stats

### Summary table

```{r}
#| message: false
#| warning: false

table_summary <-
summary_table(
  data_cleaned, 
  coordinates = coordinates[[site]], 
  location = cities[[site]], 
  site = countries[[site]], 
  color = country_colors[[site]],
  histograms = TRUE
)

table_summary

gtsave(table_summary, here("output/tables/table_summary.png"), vwidth = 850, expand = 30)
gtsave(table_summary, here("output/tables/table_summary.pdf"))
gtsave(table_summary |> cols_hide(c(plot)), here("output/tables/table_summary.docx"))
```

