---
title: "Import sleep diaries"
author: "Johannes Zauner"
format: 
  html:
    self-contained: true
    code-tools: true
---

## Preface

This document imports the `sleepdiaries` and shows descriptive statistics for the site.

## Setup

```{r}
#| message: false
library(tidyverse)
library(LightLogR)
library(glue)
library(readxl)
library(gt)
library(gtsummary)

remote <- 
  "https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/"

c("labeling",
  "radio_factors",
  "time_summaries",
  "prepare_codebook",
  "filefinder",
  "general_parameters",
  "coltype_checker",
  "tables"
) |> walk(\(x) source(paste0(remote, x, ".R")))
```

## Preparation

```{r}
#collect codebook
codebook <- prepare_codebook("MeLiDosMorningSleepDiaries_DataDictionary_2024-10-16.csv")
#collect files
files <- filefinder("sleepdiary", continuous = TRUE)
#import files
data <- read_csv(files, show_col_types = FALSE) |> drop_na(redcap_repeat_instance)
#check column types
coltype_check <- coltype_checker(codebook, data)
coltype_check$details |> gt()
#collect relevant columns: POSIXct, Date & numeric
relevant_columns <- 
  coltype_check$details |> 
  filter(expected %in% c("POSIXct", "numeric", "Date")) |> 
  pull(col) |> 
  setdiff(c("status", "scheduledate"))
#add specific character columns
relevant_columns <- c("record_id", relevant_columns, "comments")
#select relevant columns
data <- data |> select(any_of(relevant_columns))
data <- 
  data |> mutate(record_id = sprintf("KNUST_S%03d", record_id))
```

### Set sleep time zone
```{r}
site = "KNUST"
data <- 
  data |> 
  mutate(across(where(is.POSIXct), 
                \(x) force_tz(x, tzs[[site]]))
         )
```

```{r}
#label variables
data <-
data |> 
  add_radio_factors(codebook, 
                    var_col = `Variable / Field Name`, 
                    type_col = `Field Type`,
                    levels_col = `Choices, Calculations, OR Slider Labels`
                    ) |> 
  add_col_labels(codebook, var_col = `Variable / Field Name`, label_col = `Field Label`)
```

## Calculate sleep and sleepduration

```{r}
data <- 
data |> 
  rename(wake = offset, sleepprep = sleep) |> 
  mutate(sleep = sleepprep + dminutes(sleepdelay),
         sleep_duration = wake - sleep
  )

attr(data$sleep, "label") <- "Sleep onset (calculated)"
attr(data$sleep_duration, "label") <- "Sleep duration (calculated)"
```

### Ensure nighttime awake duration when there are no awakenings

```{r}
data <-
  data |> 
  mutate(
    awake_duration = replace_when(awake_duration,
                                  awakenings == 0 ~ 0)
  )
```


## Summarize results

```{r}
table_sleepdiary(data)

gtsave(table_sleepdiary(data) |> as_gt(), filename = "../output/tables/table_sleepdiary.png", vwidth = 800)
```

### Translate comments

Native language is translated into English with AI and later checked by a site researcher.

```{r}
#| label: translate into english with AI

# library(ellmer)
# 
# #Providing the relevant codebook portions
# codebook_red <-
# codebook|>
#   pmap(~ paste(paste(names(codebook), c(...), sep = ": "), collapse = ", ")) |>
#   list_c() |>
#   paste0(collapse = "newline: ")
# 
# chat <- chat_openai(paste0("Copy content and make translations according to specific instructions. They represent answers to questionnaires in a scientific field study."))
# 
# #Providing the input
# data_red <-
# data|>
#   select(record_id, comments)
# data_red <-
#   data_red |>
#   pmap(~ paste(paste(names(data_red), c(...), sep = ": "), collapse = ", "))
# 
# #creating an output structure
# type_data <- type_object(
#   record_id = type_string("use the record_id information"),
#   comments = type_string("copy the original information here", required = FALSE),
#   comments_english = type_string("translation of the original information into english", required = FALSE)
# )
# 
# data_llm <-
# parallel_chat_structured(
#   chat,
#   data_red,
#   type = type_data,
#   rpm = 500,
#   max_active = 100
# )
# 
# #Ensure that no NA is caught as string
# data_llm <-
#   data_llm |>
#   mutate(comments = case_when(comments == "NA" ~ NA, .default = comments),
#          comments_english = case_when(comments_english == "NA" ~ NA, .default = comments_english))
# 
# #check that input and output are identical
# stopifnot("Input must by identical to output check" =
#             all(data_llm$comments == data$comments, na.rm = TRUE))
# stopifnot("Input must by identical to output check" =
#             all(data_llm$record_id == data$record_id, na.rm = TRUE))
# 
# data_llm <-
#   data_llm |>
#   distinct(record_id, comments, .keep_all = TRUE)
# 
# path <- "../data/AI_translations/"
# if(!dir.exists(path)) dir.create(path, recursive = TRUE)
# write_csv(data_llm, "../data/AI_translations/sleepdiary.csv")

```

```{r}
data_llm <- 
  read_csv("../data/AI_translations/sleepdiary.csv")
#add output to original
data <- data |> left_join(data_llm, by = c("record_id", "comments"))
attr(data$comments_english, "label") <- "Comments (English translation)"
```


### Export

```{r}
data <- data |> rename(Id = record_id)
sleepdiary <- data
path <- "../data/imported/continuous/"
if(!dir.exists(path)) dir.create(path, recursive = TRUE)
save(sleepdiary, file = "../data/imported/continuous/sleepdiaries.RData")
```
