---
title: "Import sleep diaries"
author: "Johannes Zauner"
format: 
  html:
    self-contained: true
    code-tools: true
---

## Preface

This document imports the `sleepdiaries` and shows descriptive statistics for the site.

## Setup

```{r}
#| message: false
library(tidyverse)
library(LightLogR)
library(glue)
library(readxl)
library(gt)
library(gtsummary)
source("https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/labeling.R")
source("https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/radio_factors.R")
source("https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/time_summaries.R")
```

Collect the necessary external information

```{r}
codebook <- 
  read_csv(
  "https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/codebook/MeLiDosMorningSleepDiaries_DataDictionary_2024-10-16.csv", show_col_types = FALSE
  )
#clean up labels
codebook <-
  codebook |> 
    mutate(
      `Field Label` = 
        `Field Label` |> 
        str_remove_all("<div class=\"rich-text-field-label\">|<p>|<em>|</em>|</p>|</div>")
    )
```


## Collect files

The following files contain sleep-diary information.

```{r}
#path to participants
path_part1 <- "../data/raw/individual"
#path to questionnaire
path_part2 <- "/continuous/sleepdiary"
#getting all subfolders
folders <- dir(path_part1)
#creating complete folder names
paths <- glue("{path_part1}/{folders}{path_part2}")
#collecting file names
files <- list.files(paths, full.names = TRUE)
```

## Import files

```{r}
sleepdiary <- 
  read_csv(files, show_col_types = FALSE) |> drop_na(redcap_repeat_instance)
lookup_id <- 
  read_csv("../data/raw/lookup.csv", show_col_types = FALSE)
sleepdiary <- 
sleepdiary |> 
  mutate(record_id = lookup_id$ID[record_id])
```

### Check column types

```{r}
should_POSIXct <- c("bedtime", "sleep", "offset", "out_ofbed")
should_numeric <- c("sleepdelay", "awakenings", "awake_duration", "sleepquality", "daytype2")
should_character <- c("record_id", "comments")

stopifnot(
  "all of should_POSIXct need to be part of the dataset and of type POSIXct" =
  should_POSIXct %in% (sleepdiary |> select(where(is.POSIXct)) |> names()),
  "all of should_numeric need to be part of the dataset and of type numeric" =
  should_numeric %in% (sleepdiary |> select(where(is.numeric)) |> names()),
  "all of should_character need to be part of the dataset and of type character" =
  should_character %in% (sleepdiary |> select(where(is.character)) |> names())
  )
```

### Select relevant variables &nonempty rows

```{r}
sleepdiary <- 
sleepdiary |> 
  select(all_of(c(should_character, should_POSIXct, should_numeric)))
```


### Label variables

```{r}
sleepdiary <-
sleepdiary |> 
  add_radio_factors(codebook, 
                    var_col = `Variable / Field Name`, 
                    type_col = `Field Type`,
                    levels_col = `Choices, Calculations, OR Slider Labels`
                    ) |> 
  add_col_labels(codebook, var_col = `Variable / Field Name`, label_col = `Field Label`)
```

### Calculate sleep and sleepduration

```{r}
sleepdiary <- 
sleepdiary |> 
  rename(wake = offset, sleepprep = sleep) |> 
  mutate(sleep = sleepprep + dminutes(sleepdelay),
         sleep_duration = wake - sleep
  )

attr(sleepdiary$sleep, "label") <- "Sleep onset (calculated)"
attr(sleepdiary$sleep_duration, "label") <- "Sleep duration (calculated)"

```

## Summarize diary results

```{r}
table_sleep <- 
sleepdiary |> 
  tbl_summary(include = -c(comments, record_id),
              statistic = list(all_continuous() ~ "{median} ({p25}, {p75})", 
                               all_categorical() ~ "{n} ({p}%)",
                               c(bedtime, sleep, sleepprep)  ~ "{time_median} ({nighttime_p25}, {nighttime_p75})",
                               c(wake, out_ofbed)  ~ "{time_median} ({daytime_p25}, {daytime_p75})"
                               ),
              type = awakenings ~ "continuous",
              missing_text = "missing") |> 
  add_n() |> 
  bold_labels() |> 
  modify_header(label = "**Morning sleep diary**") |> 
  modify_footnote_header("Nighttime variables center on midnight, daytime variables on noon; median for time is based on circular time", columns = stat_0, replace = FALSE)

table_sleep

gtsave(table_sleep |> as_gt(), filename = "../output/tables/table_sleepdiary.png")
```

### Export

```{r}
save(sleepdiary, file = "../data/imported/continuous/sleepdiaries.RData")
```

